---
title: "Effectiveness"
author: "João Correia"
date: "11/6/2019"
output: rmarkdown::github_document
---

## Effectivenes script 

In this notebook, we will explore each step in the script [Effectiveness.R](https://github.com/correiajoao/data-analysis/blob/master/Effectiveness.R). The script explores a dataset containing software metrics and developers perception about the presence of code smells in that code or not.

### Importing pacakges

The first lines in the code, imports the required packages from the processing.

```{r}
library(RWeka)
library(e1071)
library(gmodels)
library(caret)
library(irr)
library(randomForest)
```

## Defining functions 
### Functions to calculate model's metrics

Using the results provided by confusing matrix: true positives, true negatives, false positives and false negatives; the following functions calculate individual metrics from the model's performance. The metrics are: precision, recall and f-Measure.

```{r}
# Precision
precision <- function(tp, fp){
  precision <- tp/(tp+fp)
  return(precision)
}

# Recall
recall <- function(tp, fn){
  recall <- tp/(tp+fn)
  return(recall)
}

# F-measure
f_measure <- function(tp, fp, fn){
  f_measure <- (2*precision(tp,fp)*recall(tp,fn))/(recall(tp,fn) + precision(tp, fp))
  return(f_measure)
}
```

### Functions to compute confusing matrix 

Given the predction performed by the model and the test set, the function below computes the confusing matrix and returns in array format the model's precision, recall and f-measure.

```{r}
measures <- function(test, pred){
  
  true_positive <- 0
  true_negative <- 0
  false_positive <- 0
  false_negative <- 0
  
  for(i in 1:length(pred)){
    if(test[i] == TRUE && pred[i] == TRUE){
      true_positive <- true_positive + 1
    }else if(test[i] == FALSE && pred[i] == FALSE){
      true_negative <- true_negative + 1
    }else if(test[i] == FALSE && pred[i] == TRUE){
      false_negative <- false_negative + 1
    }else if(test[i] == TRUE && pred[i] == FALSE){
      false_positive <- false_positive + 1
    }
  }
  
  measures <- c(precision(true_positive,false_positive), 
                recall(true_positive,false_negative), 
                f_measure(true_positive,false_positive,false_negative))
  
  return(measures)
}
```

### Functions to perform the machine learning algorithms

The following functions, receive as arguments the dataset and the number of folders. Using these arguments, the function split the dataset in train and test. The function build a model using the training subset, then using the same model, predicts the intances from the test subset. Aftet that, pass the predction and the test subset for the function that computes the confusing matrix and the model's metrics, returning that metrics.

The algorithms are:
  - J48
  - Naive Bayes
  - C50
  - SVM
  - One R
  - JRip
  - Random Forest
  - SMO


```{r}
executeJ48 <- function(dataset, folds){
  results <- lapply(folds, function(x) {
    train <- dataset[-x, ]
    test <- dataset[x, ]
    model <- J48(train$Smell~ ., data = train)
    pred <- predict(model, test)
    
    results <- measures(test$Smell, pred)
    
    return(results)
  })
}

executeNaiveBayes <- function(dataset, folds){
  results <- lapply(folds, function(x) {
    train <- dataset[-x, ]
    test <- dataset[x, ]
    model <- naiveBayes(train, train$Smell, laplace = 1)
    pred <- predict(model, test)
    
    results <- measures(test$Smell, pred)
    
    return(results)
  })
  
}

executeC50 <- function(dataset, folds){
  results <- lapply(folds, function(x) {
    train <- dataset[-x, ]
    test <- dataset[x, ]
    model <- C5.0(train, train$Smell)
    pred <- predict(model, test)
    
    results <- measures(test$Smell, pred)
    
    return(results)
  })
  
}

executeSVM <- function(dataset, folds){
  results <- lapply(folds, function(x) {
    train <- dataset[-x, ]
    test <- dataset[x, ]
    model <- svm(train$Smell~ ., data = train)
    pred <- predict(model, test)
    
    results <- measures(test$Smell, pred)
    
    return(results)
  })
  
}

executeOneR <- function(dataset, folds){
  results <- lapply(folds, function(x) {
    train <- dataset[-x, ]
    test <- dataset[x, ]
    model <- OneR(train$Smell~ ., data = train)
    pred <- predict(model, test)
    
    results <- measures(test$Smell, pred)
    
    return(results)
  })
  
}

executeJRip <- function(dataset, folds){
  results <- lapply(folds, function(x) {
    train <- dataset[-x, ]
    test <- dataset[x, ]
    model <- JRip(train$Smell~ ., data = train)
    pred <- predict(model, test)
    
    results <- measures(test$Smell, pred)
    
    return(results)
  })
  
}

executeRandomForest <- function(dataset, folds){
  results <- lapply(folds, function(x) {
    train <- dataset[-x, ]
    test <- dataset[x, ]
    model <- randomForest(train$Smell~ ., data = train)
    pred <- predict(model, test)
    
    results <- measures(test$Smell, pred)
    
    return(results)
  })
}

executeSMO <- function(dataset, folds){
  results <- lapply(folds, function(x) {
    train <- dataset[-x, ]
    test <- dataset[x, ]
    model <- SMO(train$Smell~ ., data = train)
    pred <- predict(model, test)
    
    results <- measures(test$Smell, pred)
    
    return(results)
  })
}
```

## Execution

### Creating data frames

In the execution, the first step is to define arrays and dataframes containing algoritms, developers and smells.

```{r}
techniques <- c("J48", "NaiveBayes", "SVM", "oneR", "JRip", "RandomForest", "SMO")
smells <- c("FE", "DCL", "GC", "II","LM", "MC", "MM", "PO","RB","SG")

developers <- data.frame(c(1, 5, 6, 9, 55, 58, 60, 84, 97, 99, 101, 103),
                         c(2, 17, 18, 19, 21, 22, 27, 30, 77, 86, 93, 104),
                         c(1, 9, 13, 15, 16, 61, 62, 66, 84, 94, 102, 103),
                         c(2, 7, 21, 22, 24, 25, 28, 86, 104, 110, 111, 124),
                         c(41, 42, 43, 45, 46, 47, 49, 51, 64, 74, 81, 95),
                         c(5, 6, 10, 52, 53, 55, 58, 60, 91, 97, 99, 101),
                         c(8, 11, 39, 40, 41, 42, 43, 45, 46, 47, 74, 81),
                         c(46, 47, 49, 51, 52, 53, 64, 74, 91, 95, 105, 109),
                         c(13, 15, 16, 17, 18, 19, 30, 61, 94, 102, 111, 112),
                         c(5, 49, 51, 52, 53, 55, 56, 64, 91, 95, 99, 105))

colnames(developers) <- smells
list_of_results <- list()

```

### Training models 

The code below interates over each developer and smell, performing the train and calculating metrics for each generated model. Each developer (total 10) has an opinion about one code smell (total 12). For each pair (developer, smell), all the seven machine learning algorithm is performed, and the metrics were stored. 

```{r}
for(j in 1:10){
  
  print(colnames(developers)[j])
  
  path <- paste("/Users/joaocorreia/data-analysis/Developers/",colnames(developers)[j],"/",colnames(developers)[j]," - ",sep="")
  
  results <- data.frame(0,0,0, 0, 0,0,0)
  
  
  for(q in 1:12){
    
    dev_path <- paste(path,developers[q,j],".csv",sep="")  
    dataset <- read.csv(dev_path, stringsAsFactors = FALSE)
    
    dataset$Smell <- factor(dataset$Smell)
    
    set.seed(3)
    folds <- createFolds(dataset$Smell, k =5)
    
    resultsJ48 <- executeJ48(dataset, folds)
    partial_results <- rowMeans(as.data.frame(resultsJ48), na.rm = TRUE)
    
    resultsNaiveBayes <- executeNaiveBayes(dataset, folds)
    partial_results <- rbind(partial_results, rowMeans(as.data.frame(resultsNaiveBayes), na.rm = TRUE) ) 
    
    resultsSVM <- executeSVM(dataset, folds)
    partial_results <- rbind(partial_results, rowMeans(as.data.frame(resultsSVM), na.rm = TRUE)) 
    
    resultsOneR <- executeOneR(dataset, folds)
    partial_results <- rbind(partial_results, rowMeans(as.data.frame(resultsOneR), na.rm = TRUE)) 
    
    resultsJRip <- executeJRip(dataset, folds)
    partial_results <- rbind(partial_results, rowMeans(as.data.frame(resultsJRip), na.rm = TRUE)) 
    
    resultsRandomForest <- executeRandomForest(dataset, folds)
    partial_results <- rbind(partial_results, rowMeans(as.data.frame(resultsRandomForest), na.rm = TRUE)) 
    
    resultsSMO <- executeSMO(dataset, folds)
    partial_results <- rbind(partial_results, rowMeans(as.data.frame(resultsSMO), na.rm = TRUE)) 
    
    rownames(partial_results) <- c("J48", "NaiveBayes", "SVM", "oneR", "JRip", "RandomForest","SMO")
    colnames(partial_results) <- c("Precision", "Recall", "F-measure")
    
    print(paste("Developer",developers[ q, j]))
    
    print(partial_results)
    
    results <- rbind(results, partial_results[,3])
  
  }
  
  results <- results[-1,]
  rownames(results) <- developers[ ,j]
  colnames(results) <- techniques
  results[,] <- lapply(results,function(x){ x[is.nan(x)]<-0;return(x)})
  
  list_of_results[[j]] <- results
  
}

```

### Print partial results

After the training and evaluation were performed, the folling print occours. That print shows a matrix containing the precision obtained for each algorithm in predicts the developer brief about code smells presence in a piece of code.

```{r}
print(list_of_results) 
```

The next line is just another way to show the same information obtained using the previous print.

```{r}
for(smell in 1:10){
  print(smells[smell])
  print(list_of_results[[smell]])
}
```

### Mean of results

Althoug the code below seems to be complex, they just calculates the mean of the results obtained by the algorithms in the prediction of conde smells, considering all developers.

```{r}
results_mean <-     matrix(c(mean(list_of_results[[1]]$J48), 
                             mean(list_of_results[[1]]$NaiveBayes), 
                             mean(list_of_results[[1]]$SVM), 
                             mean(list_of_results[[1]]$oneR), 
                             mean(list_of_results[[1]]$JRip), 
                             mean(list_of_results[[1]]$RandomForest), 
                             mean(list_of_results[[1]]$SMO)), 
                           nrow = 1,
                           ncol = 7)

for(smell in 2:10){
  results_mean <- rbind(results_mean, c(mean(list_of_results[[smell]]$J48), 
                                        mean(list_of_results[[smell]]$NaiveBayes), 
                                        mean(list_of_results[[smell]]$SVM), 
                                        mean(list_of_results[[smell]]$oneR), 
                                        mean(list_of_results[[smell]]$JRip), 
                                        mean(list_of_results[[smell]]$RandomForest), 
                                        mean(list_of_results[[smell]]$SMO)))
}
```

### Ploting graph 

Finaally the lest block of code is used to set and print a chart relating code smells and the effectiveness of the machine learning algorithms in predict them using developers opinion. 

```{r}
colnames(results_mean) <- techniques
rownames(results_mean) <- colnames(developers)
results_mean <- t(results_mean)
results_mean

barplot(results_mean, 
        main="Code Smells x Effectiveness",
        ylab="Effectiveness",
        xlab="Techniques", 
        col=c("red", "yellow", "green", "violet", "orange", "blue", "pink"), 
        ylim = c(0, 1),
        #legend = rownames(results_mean), 
        beside=TRUE)
```

